---
title: "R Notebook"
---

## Heatmap
```{r}
sce <- readRDS("/Users/reganhamel/Documents/PhD/PhD Dissertation/Chapter 5/scRNAseq/4. MG4 removed/sce.RDS")
```

```{r}
dirM <- "/Users/reganhamel/Documents/PhD/PhD Dissertation/Chapter 5/scRNAseq/4. MG4 removed/"
#sce_hm <- sce[,sce$Cluster == "MG(3B)" | sce$Cluster == "MG(3)" | sce$Cluster == "MG(3A)"]
sce_hm <- sce[,sce$Cluster == "MC(4A)"]

#sce_hm$Fabp5 <- counts(sce_hm)["Fabp5",] >= 3


# add an X to the end of the second instance of each duplicated barcode to give it a unique name (the heatmap code requires unique barcodes)
colnames(sce_hm) <- make.unique(colnames(sce_hm), sep="X")
```

#DAm1/2 genes
```{r}
source("//Users/reganhamel/Documents/PhD/scRNAseq Workflow/PhD_Dissertation/Chapter_5/Sources/Fxn - Prep Genes.R")

DAM1 <- c("Tyrobp, Ctsb, Ctsd, Apoe, B2m, Fth1, Lyz2") 

# DAM phase 2 (Trem dependent) Keren-Shaul
DAM2 <- c("Trem2, Axl, Cst7, Ctsl, Lpl, Cd9, Csf1, Ccl6, Itgax, Clec7a, Lilrb4a, Timp2")  




input <- c(DAM1, DAM2)
input <- prepG(input,sep=", ", sce)

plotHeatmap(sce_hm, features=input, exprs_values="logcounts", cellheight=5,
            zlim=c(0,9), center=FALSE, symmetric=FALSE, cluster_cols=F, cluster_rows=F,
            colour_columns_by=c("cycl","Cluster"), columns=order(sce_hm$Cluster), fontsize_row=5, fontsize_col=0.1, treeheight_col=10, file=paste0(dirM, "DAM.png"), fontsize=5, cellwidth=0.015, scale = "none")
```




#heatmap of inf vs resident in MG(3B)
#from txt
```{r}
input <- readLines(con="/Users/reganhamel/Documents/PhD/PhD Dissertation/Chapter 5/scRNAseq/3. Small Clust Removed/DEGs (some)/0.0001 & 1.25/MG3B genes.txt")

input <- c("Siglech", "Gpnmb", "Ms4a7", "Fabp5", "Lyz2", input[1:30]) 

```

```{r}
#input <- c("Fabp5", "Apoe", "Lpl", "Trem2", "Igf1", "Cst7")

plotHeatmap(sce_hm, features=input, exprs_values="logcounts", cluster_cols=F, cluster_rows=T, columns=order(sce_hm$cell_type), 
            zlim=c(0,3), colour_columns_by=c("fl", "cell_type"), file=paste0(dirM, "Mac2.png"), fontsize=3, show_rownames=T, center=F, symmetric=F, cellheight=5, cellwidth=0.1, scale = "none")

```

#mac vs microglia fig
```{r}

plotHeatmap(sce_hm, features=input, exprs_values="logcounts", cellheight=5,
            zlim=NULL, center=FALSE, symmetric=FALSE, cluster_cols=F, cluster_rows=T,
            colour_columns_by=c("fl", "cell_type", "cycl"), columns=order(sce_hm$cell_type), fontsize_row=5, fontsize_col=0.1, treeheight_col=10, file=paste0(dirM, "Mac.png"), fontsize=7, cellwidth=0.08)
```

#heatmap of neuroptrotective
```{r}
sce_hm <- sce[,grep("MG", sce$Cluster)]
sce_hm <- cbind(sce_hm, sce[,sce$Cluster == "Ctrl"])
sce_hm$Cluster <- droplevels(factor(sce_hm$Cluster))
```

#Krasemann
```{r}
#neuroprotective genes Krasemann
M0 <- c("Hexb", "P2ry12", "Mertk", "Entpd1", "Tmem119", "Tgfbr1")  #"Smad3",  not in HVG
MGnD <- c("Clec7a", 'Lgals3', "Gpnmb", "Itgax", "Spp1", "Ccr2", "Fabp5")

plotHeatmap(sce_hm, features=c(M0, MGnD), exprs_values="logcounts", cellheight=5,
            zlim=c(0,6), center=FALSE, symmetric=FALSE, cluster_cols=F, cluster_rows=F,
            colour_columns_by=c("cycl","Cluster"), columns=order(sce_hm$Cluster), fontsize_row=5, fontsize_col=0.1, treeheight_col=10, file=paste0(dirM, "neuroprotective v MGnD.png"), fontsize=5, cellwidth=0.015, scale = "none")
```

#Hif1a

cluster m1
```{r}
sce_hm <- sce[,grep("MG\\(1", sce$Cluster)]
sce_hm$Cluster <- droplevels(factor(sce_hm$Cluster))
table(sce_hm$Cluster)
```

```{r}
#neuroprotective genes Krasemann
input <- c("Hif1a","Fabp5", "Csf1r")
plotHeatmap(sce_hm, features=input, exprs_values="logcounts", cellheight=5,
            zlim=c(0,3), center=FALSE, symmetric=FALSE, cluster_cols=T, cluster_rows=F,
            colour_columns_by=c("cycl","Cluster"), columns=order(sce_hm$cycl), fontsize_row=5, fontsize_col=0.1, treeheight_col=10, file=paste0(dirM, "Hif1a.png"), fontsize=5, cellwidth=0.015, scale = "row")
```


#Cluster MG(4A) CAMs
```{r}
input  = c("Mrc1", "Lyve1", "Cd163", "Cbr2", "Ms4a7", "Cd74","H2-Eb1", "H2-Aa", "Plac8", "H2-Ab1", "H2-DMb1", "H2-DMa", "Klrd1") #Itgax (expressed in activated MG too) # not in HVG #low Siglec1 #no CAM genes in HVG


  plotHeatmap(sce_hm, features=input, exprs_values="logcounts",
            zlim=c(0,6), center=FALSE, symmetric=FALSE, cluster_cols=F, cluster_rows=T,
            colour_columns_by=c("cell_type","fl"), columns=order(sce_hm$cell_type), fontsize_row=5, fontsize_col=0.1, treeheight_col=10, file=paste0(dirM, "4A CAMs plus.png"), fontsize=5,  scale = "none", cellheight=5, cellwidth=0.2)
```

#foamy macs & LDAM
https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7217201/

```{r}
colnames(sce) <- make.unique(colnames(sce), sep="X")

sce_hm <- sce[,grep("MG", sce$Cluster)]
sce_hm <- cbind(sce_hm, sce[,grep("Ctrl", sce$Cluster)])
sce_hm$Cluster <- droplevels(factor(sce_hm$Cluster))
table(sce_hm$Cluster)
```

```{r}
library(scater)
input <- c("Fabp5","Lgals3", "Ctsb",   "Trem2", "Cd9", "Spp1", "Aldoa")
plotHeatmap(sce_hm, features=c(input), exprs_values="logcounts", cellheight=5,
            zlim=c(0,7), center=FALSE, symmetric=FALSE, cluster_cols=F, cluster_rows=F,
            colour_columns_by=c("Cluster"), columns=order(sce_hm$Cluster), fontsize_row=4, fontsize_col=0.1, treeheight_col=10, file=paste0(dirM, "Foamy LDAM MC.png"), fontsize=4, cellwidth=0.007, scale = "none")
```


#LDAM -- nothing

CD63, ATP6V1A, ATP6V1C1, ATP6V1G1 and TUBA1), genes involved in vesicular transport (RAB5B and RAB7), and CD22, a negative regulator of phagocytosis in microglia21. Interestingly, most genes linked to nitric oxide and ROS generation (for example, CAT, KL, PPP1CB, JAK and RAP1B
```{r}
LDAM <- c('Slc33a1', "Snx17", 'Vps35', 'Cln3', 'Npc2', 'Grn', 'Cd63', 'Atp6v1a', "Atp6v1c1", "Atp6v1g1", "Tuba1a", "Rab5b", "Rab7", "Cd22", "Cat", "Kl", "Ppp1cb", "Jak2", "Rap1b", "Plin3", "Acly")
LDAM <- LDAM[(LDAM %in% rownames(sce))]
```





#cargo
```{r}
colnames(sce) <- make.unique(colnames(sce), sep="X")

sce_hm <- sce
table(sce_hm$Cluster)
```

```{r}
input <- as.matrix(read.csv("/Users/reganhamel/Documents/PhD/scRNAseq Workflow/Gene Lists/Phago Cargo Clust IIb.txt"))
input <- input[input %in% rownames(sce)]
```

```{r}
library(scater)
plotHeatmap(sce_hm, features=c(input), exprs_values="logcounts", cellheight=4,
            zlim=c(0,7), center=FALSE, symmetric=FALSE, cluster_cols=F, cluster_rows=F,
            colour_columns_by=c("Cluster"), columns=order(sce_hm$Cluster), fontsize_row=4, fontsize_col=0.1, treeheight_col=10, file=paste0(dirM, "Cargo.png"), fontsize=4, cellwidth=0.01, scale = "none")
```
