---
title: "Investigating Batch effects in samples G11 and D2"
author: "Regan"
date: "01/04/2020"
output: html_document
---

# Introduction

This notebook will investigate the need for batch correction in day 1 samples

# Setup

## Load libraries

```{r}
suppressPackageStartupMessages({
  library(scater)
  library(RColorBrewer)
  library(ggplot2)
  library(scran)
  library(BiocSingular)
  source("../utils/helper_functions.R")
})
```

## Set the directory and load data

```{r}
# set path
dirH <- "/Users/reganhamel/Documents/PhD/PhD Dissertation/Chapter 3/scRNAseq/2. BC/HC DEGs/"

# load data
sceH_all <- readRDS("/Users/reganhamel/gitlab/PhD_Dissertation_data/sce.RDS")

# make column names unique
colnames(sceH_all) <- paste(colnames(sceH_all), sceH_all$sampleID, sep = "_")

# subset
sceH <- sceH_all[, sceH$stage == "HC"  | sceH_all$stage == "1" | sceH_all$stage == "2"]

```

# Investigate batch effect

### Prepare the gene lists

```{r}
cyclMG <- c("Hist1h2bc", "Birc5", "H2afx", "Ube2c", "Ccnb2", "Cenpa", "Stmn1",
            "Tuba1b", "Pclaf", "Hmgb2", 'Tubb5',"Ran", "Cks1b") #barres and Milich et al
aMG <- c("Ccl2", "Apoe", "Spp1", "Ctsb", "Lgals3", "Ftl1", "Fth1", "Fabp5",
         "Lyz2", "Mif", "Ms4a6c", "Ccl4")
hMG <- c("P2ry12", "Tmem119","Plxdc2", "Siglech", "Olfml3", "Sparc")
CAMS <- c("Mrc1", "Pf4", "Ms4a7", "Stab1", "Cbr2", "Cd36")

genes <- c(hMG, cyclMG, aMG, CAMS)

# check that they're all in the rownames
genes[!(genes %in% rownames(sceH))]
```

### Generate and save the heatmap

```{r}
name <- "CtlvAcute short_w_Milich"

plotHeatmap(sceH, features = genes, exprs_values="logcounts",
            cluster_cols = FALSE, cluster_rows = FALSE,
            columns = order(sceH$sampleID),
            colour_columns_by = c("sampleID", "stage"), fontsize = 5,
            show_rownames = TRUE, center = TRUE, symmetric = FALSE,
            scale = FALSE, cellheight = 5, cellwidth = 0.035) # remove column?
```

### Boxplot of activated genes

By sample to demonstrate that SIGAG11.D2 is activated

```{r}
genes <- c("P2ry12", "Spp1", "Mki67", "Stmn1", "Cfl1", "Birc5", "Cd36")

bxplot(sceH, x_name = "sampleID", genes = genes, colour = "stage") 
```

# Investigate Ctrl only

### Subset and normalise

```{r}
sceH <- sceH_all[, sceH_all$condition == "HC"]

subset <- "date" #all cells from a given flow cell lane will be combined into a single sce and will be batch corrected only to cells from other flow cell lanes
sceH$subset <- droplevels(factor(colData(sceH)[subset][,1]))

sceH <- batchelor::multiBatchNorm(sceH, batch = sceH$subset) #size factors are stored in sce
```

### Batch correct HCs

```{r}
# Model the variance
set.seed(100)
dec <- modelGeneVar(sceH, block = sceH$subset)

# Select the HVGs with +ve biological components to use downstream
chosen_hvgs <- getTopHVGs(dec, prop = 0.2)
str(chosen_hvgs)

sce_BC_R<- batchelor::fastMNN(sceH,
                              batch=sceH$subset,
                              subset.row = chosen_hvgs,
                              k = 20,#number of nearest neighbours considered 
                              d = 50,#number of pcs used 
                              auto.merge = TRUE, #the order of merging is important. Set to auto or order by most heterogeneity, if known
                              BSPARAM = IrlbaParam(deferred=TRUE), #set parallelization
                              correct.all = FALSE #set correct.all=T otherwise it only corrects chosen.hvg values
)

# Restore old column data and assays
colData(sce_BC_R) <- colData(sce)
counts(sce_BC_R) <- counts(sce)[HVG,]
logcounts(sce_BC_R) <- logcounts(sce)[HVG,]
sizeFactors(sce_BC_R) <- sizeFactors(sce)

# recalculate umap
set.seed(100)
sce_BC_R <- runUMAP(sce_BC_R, dimred="corrected")

# Save hvgs in metadata
metadata(sce_BC_R)$hvg <- chosen_hvgs
```

### Cluster

```{r}
set.seed(100)

metadata(sce_BC_R)$k <- 10

# Generate the shared nearest neighbour graph
# The larger k (# of nn), the fewer the clusters, 20 is the default in Seurat, 10 in scran
# Lower k means the noise will have a greater impact on the clustering.
# Larger values are computationally expensive (k=20 takes 2 mins, k=30 takes 10)
snn.gr <- buildSNNGraph(sce_BC_R, use.dimred="corrected",
                        k = metadata(sce_BC_R)$k) #k ranges from 5-100

#apply the louvain commuinity detectiono
clusters <- igraph::cluster_louvain(snn.gr) #longer random walks will lead to fewer communities

#store the clusters in the sce object column-based metadata
sce_BC_R$Cluster <- factor(clusters$membership)

# plot
col_by <- "Cluster"
plotUMAP(sce_BC_R, colour_by=col_by) #+ scale_fill_manual(values = c("#ABD9E9", "#542788"), aesthetics = "fill")

```

### Remove low quality clusters

```{r}
remove <- sce_BC_R$Cluster == 1 | sce_BC_R$Cluster == 4 | sce_BC_R$Cluster == 5 | sce_BC_R$Cluster == 6 | sce_BC_R$Cluster == 7

#get the cell barcodes
names <- colnames(sce_BC_R[,remove])
sce_full <- readRDS("/Users/reganhamel/Documents/PhD/PhD Dissertation/Chapter 3/scRNAseq/BC/sce_UMAP_noBC.RDS")

# get the index in the full sce
names_rem <- (colnames(sce_full) %in% names)

# remove and save filtered object
sce <- sce_full[,!names_rem]
saveRDS(sce, "/Users/reganhamel/Documents/PhD/PhD Dissertation/Chapter 3/scRNAseq/BC/sce_UMAP_noBC_Ctrl_filt.RDS")
```
