---
title: "Clustering"
author: "Regan"
date: "10/04/2020"
output: html_document
---

# Introduction

This notebook performs clustering on the post-batch corrected data and then annotates the cell types.

# Setup

### Load libraries

```{r}
suppressPackageStartupMessages({
  library(scater)
  library(scran)
  library(dplyr)
  library(pheatmap)
  library(tensorflow)
  library(cellassign)
})
```

### Set the directory and load data

```{r}
sce <- readRDS("/Users/reganhamel/gitlab/PhD_Dissertation_data/sce_postBC.RDS")

dirC <- "/Users/reganhamel/Documents/PhD/PhD Dissertation/Chapter 3/scRNAseq/3. Clustering/"
```

# Cluster

### Perform clustering

```{r}
set.seed(100)

metadata(sce)$k <- 20
### generate the shared nearest neighbour graph
# k ranges from 5-100
# the larger k (# of nn), the fewer the clusters, 20 is the default in Seurat, 10 in scran
# lower k means the noise will have a greater impact on the clustering. Larger values are computationally expensive 
snn.gr <- buildSNNGraph(sce, use.dimred = "corrected", k = metadata(sce)$k, type = "jaccard") 

#apply the louvain commuinity detection
clusters <- igraph::cluster_louvain(snn.gr) 

# store the clusters in the sce object column-based metadata
sce$Cluster <- factor(clusters$membership)

# preview
table(sce$Cluster)

plotUMAP(sce, colour_by = "Cluster", text_by = "Cluster")
```

### Investigate modularity between clusters

```{r}
# generating the modularity score to be used in the heatmap
cluster.mod <- clusterModularity(snn.gr, sce$Cluster, get.weights = TRUE)
log.ratio <- log2(cluster.mod$observed/cluster.mod$expected + 1)

# generating and saving the heatmap
pheatmap(log.ratio, cluster_rows = FALSE, cluster_cols = FALSE,
         color=colorRampPalette(c("white", "#7A0177"))(100))
```

# Cell type annotation

### Set the marker lists

```{r}
# lists for the cell types of interest
marker_gene_list <- list(
  Microglia = c("Sparc", "C1qa", "Plxdc2","Serpine2", "P2ry12", "Tmem119",
                "Siglech", "Ctss", "Cst3", "Slc2a5", "Sall1"), # "Sall1" not in HVG
  CAMs = c("Mrc1", "Lyve1", "Cd163", "Cbr2", "Ms4a7", "Cd74"), #low Siglec1 #no CAM genes in HVG
  Dendritic_Cells = c("Cd74", "H2-Eb1", "H2-Aa", "Plac8", "H2-Ab1", "H2-DMb1",
                      "H2-DMa", "Klrd1", "Flt3", "Zbtb46"), #Itgax (expressed in activated MG too) # not in HVG
  Monocyte = c("Chil3", "Plac8", "Ccr2", "Rgs1", "Fn1"), #Vim (expressed in others)
  Macrophage = c("Ms4a7", "Ecm1", "Arg1", "Ccl7", "Mgst1"), 
  Neutrophil = c("S100a8", "S100a9", "Lcn2", "Dedd2", "Retnlg", "Wfdc21",
                 "Mmp9", "Ly6g"), # not in HVG
  TCell = c("Cd2",   "Cd3d",  "Cd3e",  "Cd3g",  "Cd28",  "Cd4","Ptprc") #"Vim" #no TCell genes in hVGs
)


markers <- c(marker_gene_list)
marker_mat <- marker_list_to_mat(markers)

head(rownames(marker_mat))

#remove marker genes not in the HVG list
rem <- !(unlist(markers) %in% rownames(sce))
unlist(markers)[!(unlist(markers) %in% rownames(sce))]
```

### Prepare the sce with the gene list

```{r}
# subset the sce to match the marker genes
sce_marker <- sce[rownames(marker_mat)]

# keep a record of the marker genes used
saveRDS(marker_gene_list, paste0(dirA, "/marker_gene_list.RDS"))
```

### Predict the cell type

```{r}
fit <- cellassign(exprs_obj = sce_marker[rownames(marker_mat),], 
                  marker_gene_info = marker_mat, 
                  s = sizeFactors(sce_marker), 
                  shrinkage = TRUE,
                  verbose = FALSE,
                  threads = 4,
                  min_delta = 2)

# save the model
saveRDS(fit, paste0(dirA, "fit.RDS"))

#  maximum likelihood estimates (MLE)  of cell type using the `celltypes` function:
summary(factor(celltypes(fit, assign_prob = .80)))

# save the predicted cell type in the sce
sce$cell_type_ca <- celltypes(fit, assign_prob = .80)
```

### Review model

```{r}
pheatmap(cellprobs(fit))
plotUMAP(sce, colour_by = "cell_type_ca")
```

# Manual touch-ups

```{r}
#investigate the cluster membership of unasigned cells
table(sce$Cluster, sce$cell_type_ca)

sce$cell_type <- sce$cell_type_ca
```

### Manually assign cell types to unassigned cells based on cluster

```{r}
list <- list(
  "1"="Microglia",
  "2"="Macrophage",
  "3"="Monocyte",
  "4"="Neutrophil",
  "5"="Microglia",
  "6"=NULL,
  "7"="Macrophage",
  "8"="Macrophage",
  "9"=NULL,
  "10"="Microglia",
  "11"="Microglia")

for (i in 1:length(list)) {
  CT <- list[[i]]
  
  sce$cell_type[sce$Cluster == i & (sce$cell_type_ca == "unassigned" | sce$cell_type_ca == "other")] <- CT
}

#sanity check
table(sce$Cluster, sce$cell_type_ca)
table(sce$Cluster, sce$cell_type)
```

### Correct NPs

Note a few (9) NPs in the MG cluster - correct these

```{r}
sce$cell_type[sce$Cluster == 11 & sce$cell_type_ca == "Neutrophil"] <- "Microglia"
sce$cell_type[sce$Cluster == 10 & sce$cell_type_ca == "Neutrophil"] <- "Microglia"

table(sce$Cluster, sce$cell_type)
```

### Remove the NAs

```{r}
#check for NAs
sce$Cluster[is.na(sce$cell_type)]

#correct NAs based on majority in that cluster
sce$cell_type[is.na(sce$cell_type) & sce$Cluster == 2]<- 'Macrophage'
sce$cell_type[is.na(sce$cell_type) & sce$Cluster == 3]<- 'Macrophage'
sce$cell_type[is.na(sce$cell_type) & sce$Cluster == 10] <- "Microglia"
```

### Final check

```{r}
plotUMAP(sce, colour_by = "cell_type")
```

### Save

```{r}
saveRDS(sce, paste0(dirA, "sce.RDS"))
```
