---
title: "8. Misc Plots"
author: "Regan"
date: "30/06/2021"
output: html_document
---

# Introduction

This notebook generates the area composition plot, cell type vs time point plot

# Setup

### Load libraries

```{r}
suppressPackageStartupMessages({
  library(tibble)
  library(ggplot2)
  library(dplyr)
  library(RColorBrewer)
  library(corrplot)
  library(scater)
  library(scran)
  source("../utils/helper_functions.R")
})
```

### Set the directory and load data

```{r}
sce <- readRDS("/Users/reganhamel/gitlab/PhD_Dissertation_data/sce.RDS")

#reorder the cell types for aesthetics 
sce$cell_type <- factor(sce$cell_type,
                        levels = c("CAMs","TCell","Monocyte","Neutrophil","Dendritic_Cells","Macrophage","Microglia"))

dirP <- "/Users/reganhamel/Documents/PhD/PhD Dissertation/Chapter 3/scRNAseq/Misc Plots/"
```

# Generate the area plot

```{r}
sce_comp <- sce

# set the healthy control to 0
sce_comp$stage <- factor(sce_comp$stage,
                         levels = c("Ctrl","0", "1", "2", "3", "10", "21"))
sce_comp$stage[sce_comp$stage == "Ctrl"] <- 0

# set the x and y values for the plot
y <- sce_comp$cell_type
x <- droplevels(factor(sce_comp$stage))

# create the df containing the x and y values
df <- as.data.frame(table(y, x))
colnames(df) <- c("y", "x", "Freq")

# create data
time <- as.numeric(levels(df$x)[df$x]) # x Axis
value <- as.numeric(df$Freq)             # y Axis
group <- df$y       # group, one shape per group
data <- data.frame(time, value, group)

# prep data
data <- data  %>%
  group_by(time, group) %>%
  summarise(n = sum(value)) %>%
  mutate(percentage = n / sum(n))

# Plot
ggplot(data, aes(x = time, y = percentage, fill = group)) + 
  geom_area(alpha = 0.7 , size = 0.5, colour="black") + ylab("Fraction of Cells") +
  scale_x_continuous(name = "Day Post-SCI", breaks = c(0,1,2, 3, 10, 21),
                     labels = c("0"="Ctrl","1"="1","2"="2", "3"="3",
                                "10"="10", "21"="21")) +
  scale_fill_brewer(palette = "Accent", direction = -1)
```

# Histogram of cell type per cluster

```{r}
# set the x and y values for the plot
y <- sce$cell_type
x <- droplevels(factor(sce$Cluster))

# create the df containing the x and y values
df <- as.data.frame(table(y, x))
colnames(df) <- c("y", "x", "Freq")

# create data
cluster <- as.numeric(levels(df$x)[df$x]) # x Axis
value <- as.numeric(df$Freq)             # y Axis
group <- df$y       # group, one shape per group
data <- data.frame(cluster, value, group)

# prep data
data <- data  %>%
  group_by(cluster, group) %>%
  summarise(n = sum(value)) %>%
  mutate(percentage = n / sum(n))

# Plot
ggplot(data, aes(fill = group, y = value, x = cluster)) + 
  geom_bar(position = "stack", stat = "identity") +
  ylab("Number of Cells") + xlab("Cluster")  + 
  scale_x_discrete(limits = c(1:11)) +
  scale_fill_brewer(palette = "Accent", direction = -1) 
```

# Correlation of clusters between stages

### Pearson's residuals

-   <http://www.sthda.com/english/wiki/chi-square-test-of-independence-in-r> (how to interpret)

```{r}
# Create table
All <- table(sce$Cluster, sce$stage)

# Run ChiSquare test
chisq <- chisq.test(All)

# convert the z-scores (standardized residuals) to chi square values
chisq.val <- chisq$stdres * chisq$stdres

# convert the chi square values to p values
chi.p <- pchisq(chisq.val, df = 1, lower.tail = FALSE) 

# correct for multiple testing:
# compare the chi square values to the bonferonni adjusted p value
bf.p <- (0.05/(nrow(All)*ncol(All)))

# Plot
corrplot(chisq$stdres, p.mat=chi.p, is.cor = FALSE, sig.level = bf.p,
         tl.col = "black", method="color", cl.length =10, insig = "blank")
```

# Generate box plots for genes of interest

```{r}
genes <- c("Ngp", "Ftl1", "Cd52", "Ctsb", "Timp2", "Mif", "Anxa5")

bxplot(sce, x_name = "cell_type", genes = genes, colour = "cell_type") 
```

# Heatmaps

### Miscellanious gene lists

```{r}
### set the cell type
# input <- c("S100a9", "S100a8", "Retnlg", "Ifitm1", "Irg1", "Hdc", "Cxcr2", "Mxd1")
# name <- "granulocytes"

input <- c('Cd79a',  'Igll1',  'Vpreb3',  'Vpreb1',  'Lef1')
#c("Ttr", "Igkc", "Ebpp2", "Ighg2c", "Ccl5",  "Ms4a4b", "Thy1", "Icos", "Trbc1", 'Trbc2', "Gzma", "Ccl5", "Prss34", "Ly6c2", "Vim", "Mpo")
name <- "lymphocytes"

# remove any genes not in the sce object
input <- intersect(input, rownames(sce))

### plot
sce_hm <- sce

#rename the barcodes in the sce object of duplicated cell names
colnames(sce_hm) <- make.unique(colnames(sce), sep = "X")

# plot and save the heatmap
plotHeatmap(sce_hm, features = input, exprs_values = "logcounts", 
            zlim = NULL, center = FALSE, symmetric = FALSE,
            cluster_cols = FALSE,cluster_rows = TRUE,
            colour_columns_by = c("cell_type","Cluster"),
            columns = order(sce_hm$cell_type), main = paste(name),
            fontsize_row = 5, fontsize_col = 0.2, treeheight_col = 40)
```

### Non-myeloid cell genes

Prep the cell type-specific gene lists

```{r}
# load the gene lists for cell types collected from other scRNA-seq studies (see script for source articles)
source("../gene_lists/NonM_RNA.R")

# select the cell types to search for
cells <- c(Peri, Ep,Neu, Ast, Olig, MG, MC, NP, DC)
n <- c("Peri", "Ep", "Neu", "Ast", "Olig", "MG", "MC", "NP", "DC")
x <- 0
# TODO refactor this 
for (i in cells) {
  x <- x+1
  genes <- i
  input <- prepG(genes, sep = " ")
  assign(n[x], input)
}

# label the rows by cell type that the gene sets represent
rows <- data.frame(Genes = c(MG, MC, DC, NP,  Neu, Ast, Olig, Peri, Ep), 
                   Cell_Type=c(rep("Microglia", length(MG)),
                               rep("Monocyte/Monocyte-Derived", length(MC)),
                               rep("Dendritic Cells", length(DC)),
                               rep("Neutrophils", length(NP)),
                               rep("Neurons", length(Neu)),
                               rep("Astrocytes", length(Ast)),
                               rep("Oligodendrocytes_OPCs", length(Olig)),
                               rep("Pericytes", length(Peri)),
                               rep("Ependymal_Cells", length(Ep)))
)

# remove duplicated genes
rows <- rows[!duplicated(rows$Genes),] 
rownames(rows) <- as.matrix(rows$Genes)

# remove genes not in the sce
rows <- rows[(rows$Genes %in% rownames(sce_hm)),]


# Reorder and label the cell types
# add empty factors to the cell_types 
sce_hm$cell_type <- factor(sce_hm$cell_type,
                           levels = c("Microglia", "MCd", "DC", "NP", "Neurons",
                                      "Astrocytes", "Oligodendrocytes_OPCs",
                                      "Pericytes", "Ependymal_Cells"))

# add the full names of the cell types
sce_hm$cell_type <- revalue(sce_hm$cell_type, c("MG" = "Microglia",
                                                "MCd" = "Monocyte/Monocyte-Derived",
                                                "DC" = "Dendritic Cells",
                                                "NP" = "Neutrophils"))

# run with fewer cells to improve speed 
splitpro<- 0.3 ## 60/40 split
tr <- sample(ncol(sce_hm), round(ncol(sce_hm) * splitpro))
sce_hm.reduced <- sce_hm[,tr ]

# Generate and save the heatmap
plotHeatmap(sce_hm.reduced, features = rows$Genes, exprs_values = "logcounts",
            cellheight = 0.75, cellwidth = 0.01, cluster_cols = FALSE,
            cluster_rows = FALSE, columns = order(sce_hm$cell_type),
            colour_columns_by = "cell_type", annotation_row = rows,
            show_rownames = FALSE, center = FALSE, symmetric = FALSE,
            scale = "column")
```
