---
title: "8. Misc Plots"
author: "Regan"
date: "30/06/2021"
---

Area composition plot, cell type vs timepoint


Load the sce & set the directory
```{r}
sce <- sce_comp <- readRDS("/Users/reganhamel/Documents/PhD/PhD Dissertation/Chapter 4/scRNAseq/4. BC/BC by strain plus Jan/sce_postBC.RDS")


#reorder the cell types for aesthetics 
sce_comp$cell_type <- sce$cell_type <- factor(sce_comp$cell_type, levels=c("CAMs","TCell","Monocyte","Neutrophil","Dendritic_Cells","Macrophage","Microglia"))

dirP <- "/Users/reganhamel/Documents/PhD/PhD Dissertation/Chapter 3/scRNAseq/Misc Plots/"
```

Generate the area plot (using sce_comp)
```{r}
library(tibble)
library(ggplot2)
library(plyr)


# set HC to 0
sce_comp$stage <- factor(sce_comp$stage, levels = c("Ctrl","0", "1", "2", "3", "10", "21"))
sce_comp$stage[sce_comp$stage == "Ctrl"] <- 0

# set the x and y values for the plot
y <- sce_comp$cell_type
x <- droplevels(factor(sce_comp$stage))

# create the df containing the x and y values
df <- as.data.frame(table(y, x))
colnames(df) <- c("y", "x", "Freq")

# create data
time <- as.numeric(levels(df$x)[df$x]) # x Axis
value <- as.numeric(df$Freq)             # y Axis
group <- df$y       # group, one shape per group
data <- data.frame(time, value, group)


# plot
# Compute percentages with dplyr
library(dplyr)
data <- data  %>%
  group_by(time, group) %>%
  summarise(n = sum(value)) %>%
  mutate(percentage = n / sum(n))

# Plot
# Plot
ggplot(data, aes(x=time, y=percentage, fill=group)) + 
  geom_area(alpha=0.7 , size=0.5, colour="black") + ylab("Fraction of Cells") + scale_x_continuous(name ="Day Post-SCI", breaks=c(0,1,2, 3, 10, 21), labels=c("0"="Ctrl","1"="1","2"="2", "3"="3", "10"="10", "21"="21"))+ scale_fill_brewer(palette="Accent", direction=-1)


ggsave(paste0(dirP, "/Composition Area Plot cluster.png"), height=5, width=6)
```

#histogram of cell type per cluster
```{r}
library(ggplot2)
library(dplyr)


# set the x and y values for the plot
y <- sce$cell_type
x <- droplevels(factor(sce$Cluster))

# create the df containing the x and y values
df <- as.data.frame(table(y, x))
colnames(df) <- c("y", "x", "Freq")

# create data
cluster <- as.numeric(levels(df$x)[df$x]) # x Axis
value <- as.numeric(df$Freq)             # y Axis
group <- df$y       # group, one shape per group
data <- data.frame(cluster, value, group)


# plot
# Compute percentages with dplyr
data <- data  %>%
  group_by(cluster, group) %>%
  summarise(n = sum(value)) %>%
  mutate(percentage = n / sum(n))
x

# Small multiple
ggplot(data, aes(fill=group, y=value, x=cluster)) + 
  geom_bar(position="stack", stat="identity") +
  ylab("Number of Cells") + xlab("Cluster")  + 
  scale_x_discrete(limits=c(1:11)) +
  scale_fill_brewer(palette="Accent", direction=-1) 

ggsave(paste0(dirP, "/barplot_celltype_clust.png"), height=5, width=6)

```

```{r}
library(scater)
cby <- "sampleID"
plotUMAP(sce[,sce$sampleID == "SIGAG5" | sce$sampleID == "SIGAF5.D21" ], colour_by=cby, shape_by="strain")
ggsave(filename = paste0(cby, " H5_F5.png"), path = dirP, scale = 1, width = 8, height = 8,  dpi = 300, limitsize = TRUE)

```
#Correlation of clusters bt stages


#pearson's residuals
- http://www.sthda.com/english/wiki/chi-square-test-of-independence-in-r (how to interpret)
```{r}
library(RColorBrewer)
library(corrplot)

All <- table(sce$Cluster, sce$stage)

chisq <- chisq.test(All)

# convert the z-scores (standardized residuals) to chi square values
chisq.val <- chisq$stdres * chisq$stdres

# convert the chi square values to p values
chi.p <- pchisq(chisq.val, df=1, lower.tail=FALSE) 

# correct for multiple testing:
# compare the chi square values to the bonferonni adjusted p value
bf.p <- (0.05/(nrow(All)*ncol(All)))

pdf(paste0(dirP, "corrplot.pdf"))
corrplot(chisq$stdres, p.mat=chi.p, is.cor = FALSE, sig.level = bf.p, tl.col = "black", method="color", cl.length =10, insig = "blank")
dev.off()
```

Generate box plots for DEGs
```{r}
library(RColorBrewer)
library(ggplot2)
library(scran)

genes <-  "Ngp" #c("Ftl1", "Fth1", "Cd52", "Ctsb", "Timp2", "Mif", "Anxa5")
# set the clusters
x <- sce$cell_type

# generate plots for each gene
for (i in genes) {
  
  # extract the log10-transformed counts for the gene of interest
  logs <- as.numeric(logcounts(sce[i]))
  df <- data.frame(Logs=logs, x=x, cond = sce$cell_type)
  
  #plot
  d <- ggplot(data=df, aes(y=Logs, x=x, fill=cond)) + geom_boxplot() + ggtitle(i) + theme(plot.title=element_text(hjust=0.5), axis.text.x=element_text(angle=45, hjust=1)) + ylab("log10 UMI Counts") + scale_fill_manual(values=c("lightblue", brewer.pal(11,"RdYlGn"),"#003319","black", "grey50", "grey50"))  + xlab("Cluster") 
  
  # save and size
 # ggsave(paste0(dirM,"byClust -", i, ".png"), scale=1, height=3, width=4)
}

d
```

#heatmaps
#from txt
```{r}

# input <- c("S100a9", "S100a8", "Retnlg", "Ifitm1", "Irg1", "Hdc", "Cxcr2", "Mxd1")
# name <- "granulocytes"

input0 <- c('Cd79a',  'Igll1',  'Vpreb3',  'Vpreb1',  'Lef1')
  #c("Ttr", "Igkc", "Ebpp2", "Ighg2c", "Ccl5",  "Ms4a4b", "Thy1", "Icos", "Trbc1", 'Trbc2', "Gzma", "Ccl5", "Prss34", "Ly6c2", "Vim", "Mpo")
name <- "lymphocytes"

input <- match(input0, rownames(sce))
discard <- which(is.na(input))
keep <- which(!is.na(input))

input0[discard]
```

Generate the heatmaps for each cell type of interest 
```{r}
library(scater)

# set the cell type

sce_hm <- sce

#rename the barcodes in the sce object of duplicated cell names
colnames(sce_hm) <- make.unique(colnames(sce), sep = "X")


# plot and save the heatmap
plotHeatmap(sce_hm, features=input, exprs_values="logcounts", 
            zlim=NULL, center=FALSE, symmetric=FALSE, cluster_cols=FALSE,cluster_rows=T,
            colour_columns_by=c("cell_type","Cluster"), columns=order(sce_hm$cell_type), main= paste(name), fontsize_row=5, fontsize_col=0.2, treeheight_col=40, file=paste0(dirP, name, ".png"))  #cellheight=4, cellwidth=0.03,
```


# Heatmaps of non-myeloid cell genes
Load and prep the sce
```{r}
library(scater)
sce.hm <- readRDS("/Users/reganhamel/Documents/PhD/PhD Dissertation/Chapter 4/scRNAseq/4. BC/BC by strain plus Jan/sce_postBC.RDS")
dirH <- "/Users/reganhamel/Documents/PhD/PhD Dissertation/Chapter 4/scRNAseq/Misc Plots/"

# add an X to the end of the second instance of each duplicated barcode to give it a unique name (the heatmap code requires unique barcodes)
names <- make.unique(colnames(sce.hm), sep="X")

# rename the barcodes in the sce object of duplicated cell names
colnames(sce.hm) <- names
```

Prep the cell type-specific gene lists
```{r}
# load the gene lists for cell types collected from other scRNA-seq studies (see script for source articles)
source("//Users/reganhamel/Documents/PhD/scRNAseq Workflow/PhD_Dissertation/Chapter_4/Sources/Non-M RNA.R")

#load the function to prep the gene lists
source("/Users/reganhamel/Documents/PhD/scRNAseq Workflow/PhD_Dissertation/Chapter_4/Sources/Fxn - Prep Genes.R")

# select the cell types to search for
cells <- c(Peri, Ep,Neu, Ast, Olig, MG, MC, NP, DC)
n <- c("Peri", "Ep", "Neu", "Ast", "Olig", "MG", "MC", "NP", "DC")
x <- 0
for (i in cells) {
  x <- x+1
  genes <- i
  input <- prepG(genes, sep = " ")
  assign(n[x], input)
}

# label the rows by cell type that the gene sets represent
rows <- data.frame(Genes=c(MG, MC, DC, NP,  Neu, Ast, Olig, Peri, Ep), 
                   Cell_Type=c(rep("Microglia", length(MG)),  rep("Monocyte/Monocyte-Derived", length(MC)),  rep("Dendritic Cells", length(DC)),rep("Neutrophils", length(NP)), rep("Neurons", length(Neu)), rep("Astrocytes", length(Ast)), rep("Oligodendrocytes_OPCs", length(Olig)),  rep("Pericytes", length(Peri)), rep("Ependymal_Cells", length(Ep))))

# remove duplicated genes
rows <- rows[!duplicated(rows$Genes),] 
rownames(rows) <- as.matrix(rows$Genes)

# remove genes not in the sce
rows <- rows[(rows$Genes %in% rownames(sce.hm)),]
```

Reorder and label the cell types
```{r}
library(plyr)
library(RColorBrewer)

# add empty factors to the cell_types 
sce.hm$cell_type <- factor(sce.hm$cell_type, levels=c("Microglia", "MCd", "DC", "NP", "Neurons", "Astrocytes", "Oligodendrocytes_OPCs", "Pericytes", "Ependymal_Cells"))

# add the full names of the cell types
sce.hm$cell_type <- revalue(sce.hm$cell_type, c("MG" ="Microglia", "MCd"="Monocyte/Monocyte-Derived", "DC"="Dendritic Cells", "NP"="Neutrophils"))
```

#run less data
```{r}
#sce.all <- sce

splitpro<- 0.3 ## 60/40 split
tr <- sample(ncol(sce.hm), round(ncol(sce.hm) * splitpro))
sce.hm. <- sce.hm[,tr ]
```

Generate and save the heatmap
```{r}
sce.hm <- sce.hm.
plotHeatmap(sce.hm, features=rows$Genes, exprs_values="logcounts", cellheight=.75, cellwidth=0.01, cluster_cols=F, cluster_rows=F, columns=order(sce.hm$cell_type), colour_columns_by="cell_type", file=paste0(dirH, "Extended Data Fig.1 CNS Cells.png"), fontsize=7, annotation_row=rows, show_rownames=F, center=F, symmetric=F, scale="column")

#save as png and as pdf
```
